---
title: "Exploring the Impact of Atypical Events on Regret: Analysis Through Hitchhiker, Car Accident, and Robbery Scenarios"

author: Dailin(Ben) li
thanks: "Code and data are available at: https://github.com/UofT-DailinLi/NHL-Florida-Panthers-Team-Analysis.git" 
date: today

editor: visual

toc: true
toc_depth: 2
number-sections: true
bibliography: references.bib
format: pdf

---

\newpage

```{r echo=FALSE, message=FALSE}
#install.packages("stargazer")

# Load libraries
library(tidyverse)
library(stargazer)
library(kableExtra)
library(here)
library(dplyr)




```

# Introduction
The Florida Panthers were chosen for analysis because it was a unique opportunity to explore the dynamics of an ice hockey team in a southern city. Typically, ice hockey is played in cold areas, such as Canada and the United States’ northern cities. Ice culture has popularized the sport. However, the Panthers hockey team's location in the Miami metropolitan area is known for its warm climate. This is where we can look at a particular case to see how franchises have evolved and dealt with the challenges in a non-traditional field hockey market.


# Part 1: Florida Panthers
Although the Panthers are located in a southern city, the Miami metropolitan area actually has a significant population of 62 million, which is seventh in the U.S. Metropolitan Statistical Area (MSA) rankings[@NHLCitiesPopulation]. But hockey isn't as popular as traditional sports like football and basketball in southern cities, and even with a large market size, market penetration and fan engagement can be significantly reduced. Income levels in South Florida vary widely; as we can see from point2home's online data, the average household income is $79,886. However, by region, the highest average household income is $161,979, and the lowest is $35,837[@Miamiincomes2023]. Higher incomes in certain areas set the stage for revenue from pricey season tickets and front-row tickets. Also, the South Florida area is home to many corporate headquarters, including as many as five Fortune 500 companies. This lays a strong foundation for corporate sponsorships and partnerships, such as Ford being a major sponsor of the Panthers, which is a significant part of the team's revenue[@FloridaPanthersForbes].

Regarding team revenue, ticket revenue is the most intuitive revenue, and the Panthers have $49 million in ticket revenue based on 2023 numbers, compared to his total revenue of $160 million, or about 30 percent. And the top teams in the rankings can all have ticket revenue of about 50% of their total revenue. This revenue is affected by the team's record, publicity, and whether or not a recognizable player is competing, and the Panthers have a lot of room for improvement. The remaining 70% of revenue is from television broadcasts, corporate sponsorships, and franchise merchandise sales[@FloridaPanthersForbes]. National telecast deals are shared among NHL teams, but local broadcast deals are based on market size and team popularity. Meanwhile, the Panthers have established strong relationships with many major corporations headquartered in Miami, including arena naming rights, advertising, and sponsorships that are a significant part of the team's revenue. Team-branded merchandise and arena food sales are additional major sources of revenue.


In terms of expenses, player salaries are the most significant cost, with player expenses up to 90 million in 2023, which is 56% of total revenue, a percentage with expenses higher than many of the top few big teams. This could be related to the team's popularity and location, where the team has to pay high prices to retain talent. Again, most of the expenses are game operations, arena maintenance, and employee expenses. In 2023, the Panthers had a net income of 6 million, a vast improvement from the opposing net income before the epidemic[@FloridaPanthersForbes]. The team's dynamic and diverse revenue streams have allowed the Panthers to turn around their profit and loss, again adapting to the unique market conditions of South Florida.

# Part 2: League analysis

## Data
To run further league analysis and investigate why the Florida Panthers, a southern city-based hockey team, can survive in a complex sports environment, I gathered data from various sources, including Professor Rodney Fort's website[@RodsCodes], Forbes[@Forbes], Statistics Canada[@StatsCan], and the United States Census Bureau. I collected NHL team data from 2014 to 2019 for enough observations to run a regression analysis. It includes 182 observations and nine original variables: "Year," "Team," "Revenue," "Expense," "Payroll," "Operating Income," "NHL Points," "NHL Win Per," and "Population." To account for the inflation, I will adjust four variables, "Revenue," "Expense," "Payroll," and "Operating Income," by inflation rate to ensure they reflect constant dollars with the base year 2014, providing more accurate analysis and financial trends[@MinneapolisFedInflationCalculator]. Data was cleaned and analyzed using the open-source statistical programming language R [@r], and additional packages from `tidyverse` [@tidy], `stargazer` [@stargazer], `kableExtra` [@kableExtra], `here`[@here], and `dplyr` [@dplyr]. The cleaned dataset contains 182 observations, and nine cleaned variables will be shown below.


```{r echo=FALSE, message=FALSE, warning=FALSE, label="tbl-cleaned"}
#| tbl-cap: "First Ten Rows of Cleaned Sports Data"


sports_data = read_csv(
    file = here("inputs/sport_Data.csv"), show_col_types = FALSE)

clean_data <-sports_data %>% select(1:2, (ncol(sports_data)-7):ncol(sports_data))
colnames(clean_data)[4] <- "Winning_Percent"


# Remove rows with any NA values
df_clean <- na.omit(clean_data)
# Convert Population to millions
df_clean$Population <- df_clean$Population / 1e6




# write.csv(df_clean, here("outputs/data/cleaned_Sports_Data.csv"), row.names = FALSE)






kable(head(df_clean, 10), "latex", booktabs = T) %>%
  kable_styling(font_size = 13, latex_options = c("scale_down", "hold_position")) 

```

@tbl-cleaned shows the first ten rows of cleaned sports data, "years" means the data was estimated in which year, "Team" means the sports team, "NHL Points" refers to a scoring system used in the NHL league, 2 points for a win, 1 point for an overtime or shootout loss, 0 points for a loss in regulation time. "Winning Percent" means the win rate of a team, and "population" means the Combined Statistical Area (CSA) population. The "Adjusted Revenue," "Adjusted Exp," "Adjusted Payroll (Million)," and "Adjusted Income" are four adjusted income-related variables.

## Regressiin Model


```{r echo=FALSE, message=FALSE}
# Convert Year to a factor
df_clean$Year <- as.factor(df_clean$Year)
model_1 <- lm(Adjusted_Rev ~ Population, data = df_clean)
model_2 <- lm(Adjusted_Rev ~ Population +  Winning_Percent  + Year, data = df_clean)
model_3 <- lm(Adjusted_Rev ~ Population + Adjusted_Payroll_million + Year +  Winning_Percent, data = df_clean)
```

```{r echo=FALSE, message=FALSE}
stargazer(model_1, model_2, model_3, type = "text",
          title="Comparison of Linear Regression Models",
          header = FALSE, # Adjust based on your output preference
          model.names = TRUE,
          intercept.bottom = FALSE, # Puts intercept at the top
          
          font.size = "small",
          column.sep.width = "3pt" # Adjusts text for better readability 
          )
```

















# Part 3: Team in the context of the league






\newpage
# Reference